Trending assessment for nowcasts, measurements, and forecasts

### 2024-03-18 paper.tex
Update on Overleaf.

\begin{abstract}
An intuitive and easily accessible information on a quantity measured over time is whether it increases or decreases.
Trending refers to the ability to predict the direction of change for a true variable accurately. 
Thus, an appropriate assessment of the trending ability is crucial for measurements, nowcasts, and forecasts for decision-makers. 
While the current practice of evaluating signals is based on averaging over time-step-separated measures, we review and introduce trending measures that consider whether the direction of development fits the quantity of interest. 
We formalize the concept of trending for all application areas and quantify the trending ability for signals. 
We review and present new visualization techniques to assess the local trending behaviors and quantify estimation uncertainty by bootstrapping. 
We apply the methods to three large data applications: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
\end{abstract}

>>>>

\begin{abstract}
An intuitive and easily accessible information on a quantity measured over time is whether it increases or decreases.
Trending refers to the ability to predict the direction of change for a true variable accurately. 
Thus, an appropriate assessment of the trending ability is crucial for measurements, nowcasts, and forecasts for decision-makers. 
While the current practice of evaluating signals is based on averaging over local deviation measures, we review and introduce trending measures that consider whether the direction of development fits the quantity of interest. 
We formalize the concept of trending for all application areas and present quantitative measures. 
Using bootstrap confidence intervals and exclusion areas, we provide extensions for noisy data and estimation uncertainty.
We review and present new visualization techniques to assess the local trending behaviors and enlarge the scope to probabilistic nowcasts and forecasts. 
We apply the methods to three large data applications: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
\end{abstract}

### 2024-03-18 sections/01-introduction.tex
Update on Overleaf.

The evaluation of measurement, prediction, and forecasting methods become increasingly important and sophisticated as technologies and data availability enable their application in more and more areas. 
Conventionally, methods are evaluated using distance measures that take aggregated local absolute differences between predictions (or measurements) and target values into account. 
These measures fail to consider whether the right direction of change is predicted or measured correctly.
The information about whether a signal, be it a measurement device, forecast, or nowcast, can accurately predict an increase or decrease is crucial when making decisions based on the estimator's prediction results. 
In the following, we provide an overview over the characteristics and current evaluation schemes of the three application areas, measurement, forecasting, and nowcasting methods.
We highlight why trending evaluation is highly relevant in the respective fields.

Measurement aims to obtain accurate and reliable data about the current state of a system. 
A parameter or variable is can be measured regularly over a certain period to evaluate the systems development.
When introducing new measurement methods they are evaluated against current measurement techniques, called gold standard, by measuring the same quantity in different settings, for example, different individuals or environments.
In addition to using a paired t-test to determine systematic differences between two measurement series (\cite{watson2010method}), various indices such as the interclass correlation coefficient or Lin's concordance correlation have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline}). 
Furthermore, graphical tools were developed, such as the Bland-Altman diagram, visualizing the differences between the methods in relation to their mean value (\cite{bland1986statistical}). 
Trending considerations, that is the consistency of the test methods development with the gold standard, have already gained importance in literature (\cite{Saugel2015,saugel2018error,hiraishi2021concordance}). 
In this paper, we build upon the existing methods and extend them with new functionality.

>>>>

The evaluation of measurement, prediction, and forecasting methods becomes increasingly important and sophisticated as technologies and data availability enable their application in more and more fields. 
Conventionally, methods are evaluated using distance measures of local differences between predictions (or measurements) and target values into account. 
These measures fail to consider whether the right direction of change is predicted or measured correctly.
The information about whether an increase or decrease is predicted correctly is crucial when making decisions based on the estimator's prediction results. 
In the following, we provide an overview of the characteristics and current evaluation schemes of the three application areas: measurement, forecasting, and nowcasting.
We highlight why trending evaluation is highly relevant in the respective fields.

Measurement aims to obtain accurate and reliable data about the current state of a system. 
A parameter or variable can be measured regularly over a certain period to evaluate the system's development.
When introducing new measurement methods, they are evaluated against current measurement techniques, called the gold standard, by measuring the same quantity in different settings, such as individuals or environments.
In addition to using a paired t-test to determine systematic differences between two measurement series (\cite{watson2010method}), various indices such as the interclass correlation coefficient or Lin's concordance correlation have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline}). 
Furthermore, graphical tools were developed, such as the Bland-Altman diagram, visualizing the differences between the methods in relation to their mean value (\cite{bland1986statistical}). 
Trending considerations, that is, the consistency of the test methods development with the gold standard, have already gained importance in literature (\cite{Saugel2015,saugel2018error,hiraishi2021concordance}). 
In this paper, we build upon the existing methods and extend them with new functionality.
